---
title: "Legends"
author: "JB"
toc: true
number-sections: true
format: 
  html:
    embed-resources: true
---
# Basic Box Plot
- `library(ggplot2)`: This loads the ggplot2 library.
- `ggplot(PlantGrowth, aes(x = group, y = weight))`:
  - Uses the built-in `PlantGrowth` dataset
  - Creates a plot with `group` on the x-axis and `weight` on the y-axis
  - `aes()` defines the aesthetic mappings of the plot
- `geom_boxplot`()`: Adds a box plot geometry to visualize the distribution of weights across different groups
- `ggtitle('Basic Box Plot')`: Adds a title to the plot
- `theme(plot.title = element_text(hjust=0.5))`: Centers the title horizontally `(hjust=0.5)`

```{r}
library(ggplot2)
ggplot(PlantGrowth, aes(x = group, y = weight)) + 
  geom_boxplot() +
  labs(title = 'Basic Box Plot', caption = 'JB, DV, THU 2024') +
  theme(plot.title = element_text(hjust=0.5))
```

# Add Legend
By adding `fill` or `color` options to `aes()`, a legend is created automatically on the right side of the plot.

- `library(gridExtra)`: This loads the `gridExtra` library to use `grid.arrange`
- `fill`: Fill in the `colors` inside the chart
- `color`: Change the `color` of the outlines of the chart
- `grid.arrange`: Arrange the order of the chart
- `top`: Add `text` on the top of the chart
- `bottom`: Add `text` on the bottom of the chart


```{r}
library(gridExtra)

p1 <- ggplot(PlantGrowth, aes(x = group, y = weight)) + 
  geom_boxplot()
p2 <- ggplot(PlantGrowth, aes(x = group, y = weight, fill = group)) + 
  geom_boxplot()
p3 <- ggplot(PlantGrowth, aes(x = group, y = weight, color = group)) + 
  geom_boxplot()
grid.arrange(p1,p2,p3,ncol=3,top='Add Legend', bottom = 'JB, DV, THU 2024')
```

